---
title: "gfw-spire-only"
author: "Ben Best"
date: "6/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GFW Spire only

Source: world-fishing-827:pipe_production_v20201001.messages_scored_ 
* Satellite Timing: world-fishing-827:gfw_research.pipe_v20201001_satellite_timing 
* Sunrise: world-fishing-827:pipe_static.sunrise 
* Norad to Receiver: world-fishing-827:pipe_static.norad_to_receiver_v20200127 
* Satellite positions one second resolution: world-fishing-827:satellite_positions_v20190208.satellite_positions_one_second_resolution_ 

```{r cars}
source(here::here("scripts/db.R"))
glue <- glue::glue

# messages_scored_table                     <- "world-fishing-827.pipe_production_v20201001.messages_scored_"
messages_scored_table                     <- "world-fishing-827.pipe_production_v20201001.messages_scored_"
research_satellite_timing_table           <- "world-fishing-827.gfw_research.pipe_v20201001_satellite_timing"
static_sunrise_dataset_and_table          <- "world-fishing-827.pipe_static.sunrise"
static_norad_to_receiver                  <- "world-fishing-827.pipe_static.norad_to_receiver_v20200127"
satellite_positions_one_second_resolution <- "world-fishing-827.satellite_positions_v20190208.satellite_positions_one_second_resolution_"
date <- Sys.Date() - 7

sql <- readLines("gfw/research_daily.sql") %>% 
  paste(collapse = "\n") %>% 
  glue() # cat(sql)

(projects <- bq_projects())
prj <- projects[[1]]

tb <- try(bq_project_query(prj, sql))

if (!"try-error" %in% class(tb))
   try(bq_table_download(tb, max_results = 10))
```

```
Error : Job 'benioff-ocean-initiative.job_y8EvokYB59cNQ-TQt4zf8tVdcFuD.US' failed
x Access Denied: Table world-fishing-827:gfw_research.pipe_v20201001_satellite_timing: User does not have permission to query table world-fishing-827:gfw_research.pipe_v20201001_satellite_timing. [accessDenied]
```

```{r}
tbls <- list(
  messages_scored_table = glue("
    SELECT * FROM `{ messages_scored_table }*` WHERE _TABLE_SUFFIX = '20210622'"),
  research_satellite_timing_table = 
    glue("SELECT * FROM `{ research_satellite_timing_table }`"),
  # x Access Denied: Table world-fishing-827:gfw_research.pipe_v20201001_satellite_timing: User does not have permission to query table world-fishing-827:gfw_research.pipe_v20201001_satellite_timing. [accessDenied]
  static_sunrise_dataset_and_table = glue("
    SELECT * FROM `{ static_sunrise_dataset_and_table }`"), # denied
  static_norad_to_receiver = glue("
    SELECT * FROM `{ static_norad_to_receiver }`"),
  satellite_positions_one_second_resolution = glue("
    SELECT * FROM `{ satellite_positions_one_second_resolution }*` WHERE _TABLE_SUFFIX = '20210622'"))

for (tbl in names(tbls)){ # tbl = names(tbls)[3]
  table <- get(tbl)
  sql   <- glue("{tbls[[tbl]]} LIMIT 2")
  message(glue("{tbl}: {get(tbl)}\n\n{sql}\n", .trim = F))
  
  tb <- try(bq_project_query(prj,sql))

  if (!"try-error" %in% class(tb))
     try(bq_table_download(tb))
}

```


